Click chemistry reveals an alternate mechanism for HPII catalase

### 2018-02-05 bb494/manuscript.tex
all

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional packages needed here.  Only include packages
%% which are essential, to avoid problems later. Do NOT use any
%% packages which require e-TeX (for example etoolbox): the e-TeX
%% extensions are not currently available on the ACS conversion
%% servers.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Meta-data block
%% ---------------
%% Each author should be given as a separate \author command.
%%
%% Corresponding authors should have an e-mail given after the author
%% name as an \email command. Phone and fax numbers can be given
%% using \phone and \fax, respectively; this information is optional.
%%
%% The affiliation of authors is given after the authors; each
%% \affiliation command applies to all preceding authors not already
%% assigned an affiliation.
%%
%% The affiliation takes an option argument for the short name.  This
%% will typically be something like "University of Somewhere".
%%
%% The \altaffiliation macro should be used for new address, etc.
%% On the other hand, \alsoaffiliation is used on a per author basis
%% when authors are associated with multiple institutions.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Elliott Capek}
\altaffiliation{Department of Biochemistry, Oregon State University}
%% \author{}
%% \altaffiliation{}
%% \author{}
%% \altaffiliation{}
%% \email{}
%% \phone{}
%% \fax{}
%% \affiliation[]{}
%% \alsoaffiliation[]{}
%% \author{}
%% \email{}
%% \affiliation[]{}
\author{Aidan Estelle}
\affiliation[]{}
\alsoaffiliation[]{}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Some journals require a list of abbreviations or keywords to be
%% supplied. These should be set up here, and will be printed after
%% the title and author information, if needed.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\abbreviations{IR,NMR,UV}
\keywords{American Chemical Society, \LaTeX}

@@

  %% This is an abstract for the \textsf{achemso} document class
  %% demonstration document.  An abstract is only allowed for certain
  %% manuscript types.  The selection of \texttt{journal} and
  %% \texttt{manuscript} will determine if an abstract is valid.  If
  %% not, the class will issue an appropriate error.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

\section{Methods}
\textbf{Protein expression}\\
Wildtype protein was expressed by transforming the pBAD-HPII-WT plasmid in Figure \ref{fig:pbad-hpii-wt} into competent \textit{E. coli} DH10B cells via heat-shock. Plasmid includes ampicillin resistance and a 6xHis tag on the N-term of the HPII gene. Cells were then grown in autoinduction media with an arabinose inducer (0.1\% aspartate pH 7.5, 0.2\% glycerol, mineral salts, 0.05\% glucose, 20mM MgSO4, 0.05\% arabinose, trace metals and AA mix) for 48 hours, shaking, at 37C. Cells were then pelletized and stored at -80C.\\

@@

\section{Results and discussion}

\begin{figure}
\centering
\includegraphics[width=\linewidth]{figures/crude-wt-gel}
\caption{Electrophoresis gel of wildtype expression. Serial dillution of WT protein purification shown in two lanes. Major band in purification lanes correspond to protein of approximately 75kD, as expected for catalse single subunit.}
\label{fig:duration-length}
\end{figure}

\subsection{Outline}

@@

\cite{Mena2000,Abernethy2003,Friedman-Hill2003,EuropeanCommission2008}.

\bibnote{This is a note. The text will be moved the the references section.  The title of the section will change to ``Notes and References''.}

\textsf{natbib}

\citeauthor{Abernethy2003}
\textsf{natbib}
\citeyear{Cotton1999}
~\citenum{Mena2000}
\cite{Pople2003}
\textsf{natbib}
\texttt{achemso-demo.bib}
\latin{et al.}
\texttt{\textbackslash latin}
\textsf{mciteplus}
\cite{Johnson1972,*Arduengo1992,*Eisenstein2005,*Arduengo1994}
\texttt{\textbackslash mciteSubRef}
\mciteSubRef{Eisenstein2005}
\cite[p.~1]{Cotton1999}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional packages needed here.  Only include packages
%% which are essential, to avoid problems later. Do NOT use any
%% packages which require e-TeX (for example etoolbox): the e-TeX
%% extensions are not currently available on the ACS conversion
%% servers.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
\usepackage{amsmath}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Meta-data block
%% ---------------
%% Each author should be given as a separate \author command.
%%
%% Corresponding authors should have an e-mail given after the author
%% name as an \email command. Phone and fax numbers can be given
%% using \phone and \fax, respectively; this information is optional.
%%
%% The affiliation of authors is given after the authors; each
%% \affiliation command applies to all preceding authors not already
%% assigned an affiliation.
%%
%% The affiliation takes an option argument for the short name.  This
%% will typically be something like "University of Somewhere".
%%
%% The \altaffiliation macro should be used for new address, etc.
%% On the other hand, \alsoaffiliation is used on a per author basis
%% when authors are associated with multiple institutions.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Elliott Capek}
\affiliation{Department of Biochemistry, Oregon State University}
\email{capeke@oregonstate.edu}
\phone{(971) 207 8193}
%% \alsoaffiliation[]{}
\author{Aidan Estelle}
%% \affiliation{Department of Biochemistry, Oregon State University}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Some journals require a list of abbreviations or keywords to be
%% supplied. These should be set up here, and will be printed after
%% the title and author information, if needed.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\keywords{Catalase, noncanonical amino acid, active site, channel flow}

@@

  %% This is an abstract for the \textsf{achemso} document class
  %% demonstration document.  An abstract is only allowed for certain
  %% manuscript types.  The selection of \texttt{journal} and
  %% \texttt{manuscript} will determine if an abstract is valid.  If
  %% not, the class will issue an appropriate error.
  Abstract abstract abstracts abstract. Abstract abstract abstracts abstract. Abstract abstract abstracts abstract. Abstract abstract abstracts abstract. Abstract abstract abstracts abstract. Abstract abstract abstracts abstract. Abstract abstract abstracts abstract. Abstract abstract abstracts abstract.\\
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Catalase is an enzyme which catalyzes the reaction $2H_2O_2 \rightarrow O_2 + 2H_2O$. This is an important detoxification reaction, since hydrogen peroxide is a reactive toxin generated by several biological processes.

The molecular mechaninism of catalase proceeds in two steps. First, the heme reduces a hydrogen peroxide, producing water and the so-called Cpd I. Then Cpd I oxidizes a second hydrogen peroxide, producing water and oxygen gas. This second step is mediated by a histidine near the active site \cite{alfonso-prieto}.

This bimolecular mechanism relies on the efficient transport of hydrogen peroxide into the deeply buried active site. This is believed to be aided by an electrical gradient over 12 \AA between an active site aspartic acid and the positive heme iron \cite{chelikani}. Mutagenisis of this Asp to positive residues decreases activity up to 90\% \cite{chelikani}. However, this study was accomplished using canonical amino acids, and provides relatively weak evidence for the electric field hypothesis.\\

In this study we seek to use the technique of bioorthogonal ligation to study the electric field hypothesis of catalase substrate transport. We will incorporate the noncanonical amino acid 3-nitrotyrosine into the active site, then use click chemistry to insert a gradient of charged particles into the active site. By using a large number of particles with charges varying from positive to neutral to negative, we can average out the effects of steric bulk and isolate the electrical effects of charged amino acids in the catalase active site. The results of such a study would conclusively determine the importance of an electric field in the catalase active site.\\

\section{Methods}
\textbf{Protein expression}\\
Wildtype protein was expressed by transforming the plasmid pBad-HPII (Addgene ID 105839), containing WT HPII gene, into competent \textit{E. coli} DH10B cells via heat-shock. pBad-HPII plasmid includes ampicillin resistance, an imidazole promoter, and a 6xHis tag on the N-term of the HPII gene. Cells were then grown in autoinduction media with an arabinose inducer (0.1\% aspartate pH 7.5, 0.2\% glycerol, 100$\mu g / mL$ ampicillin, mineral salts, 0.05\% glucose, 20mM MgSO4, 0.05\% arabinose, trace metals and AA mix) for 48 hours, shaking, at 37C. Cells were then pelletized and stored at -80C.\\

The same protocol was followed for the ncAA mutant catalases F206NY and F206pAzF, except protein plasmid pBad-HPII TAG206 (Addgene ID 105846) was used as a base. ncAA translation machinery plasmids pDule-3-nitrotyrosine (Addgene ID 85498) and pDule-pCNF (85494), which contain a tetracycline resistance gene, a TAG tRNA, and a compatible tRNA synthase, were co-transformed into DH10B along with pBAD-HPII TAG206 for respective F206NY and F206pAzF cell lines. ncAA autoinduction media was supplemented with $25\mu g / mL$ tetracycline to select for both plasmids.\\

OD600 measurements of cultures at 48h growth time were 0.65, 0.58 and 0.45 arb. units for WT, F206NY and F206pAzF cells.\\

@@

\textbf{Bradford assay}\\
A Bio-Rad Bradford assay kit was used using the standard protocol. Bovine serum albumin was used as the reference protein to construct the standard curve. Concentrations from 125 $\mu g / mL$ to 2 mg/mL were explored. Samples were incubated with dye in cuvettes for 5 minutes at RT. Samples were then analyzed via spectrophotometer at 595 \textit{nm} and used to create the standard curve shown in Figure \ref{fig:bradford-assay}. Protein samples were then compared to this standard curve to estimate their concentration.\\

\textbf{Qualitative activity assay}\\
10 $\mu g$ purified WT catalase was added to a 5\% solution of hydrogen peroxide in water. The mixture was then examined for bubble formation.\\

\section{Results}
DH10B cell lines with appropriate plasmids were grown for 48h and pelletized. Wildtype HPII-containing cells were purified using TALON resin. Purified wildtype protein and ncAA cell lysates are shown in Figure \ref{fig:crude-wt-gel}. The expected band corresponding to an 84-kD protein is seen across all lanes, verifying the presence of HPII wildtype in all cultures.\\

A Bradford assay of wildtype purified protein was then conducted. The results, shown in Figure \ref{fig:bradford-assay}, further verify the presence of protein in the purified sample.\\

A qualitative, rudimentary assay testing the activity of wildtype purified HPII catalase was conducted. This assay verified HPII-WT was active by producing bubbles.\\

\begin{figure}
\centering
\includegraphics[width=\linewidth]{figures/crude-wt-gel}
\caption{Electrophoresis gel of wildtype expression. Serial dillution of WT protein purification shown in lanes three and four. Major band in purification lanes correspond to protein of approximately 80kD, as expected for a catalase single subunit. }
\label{fig:crude-wt-gel}
\end{figure}

\begin{figure}
  \centering
  \begin{minipage}{.4\textwidth}
    \begin{tabular}{lll}
      \hline
      Sample  & OD590 & mg/mL  \\
      \hline
      WT        & 2.18 & 2.74\\
      %% F206NY    & 0.71 & 0.83\\
      %% F206pAzF  & 1.51 & 1.88\\
      \hline
    \end{tabular}
  \end{minipage}
  \begin{minipage}{.59\textwidth}
    \includegraphics[width=\linewidth]{figures/bradford-standard-curve}
  \end{minipage}
  \caption{Bradford assay results. \textbf{Left:} Sample absorbance at 590 nm and extrapolated concentration. \textbf{Right:} Bradford standard curve.}
  \label{fig:bradford-assay}
\end{figure}

@@
